---
title: "Equal Pay Analysis"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```

# Omega Group plc- Pay Discrimination


At the last board meeting of Omega Group Plc., the headquarters of a large multinational company, the issue was raised that women were being discriminated in the company, in the sense that the salaries were not the same for male and female executives. A quick analysis of a sample of 50 employees (of which 24 men and 26 women) revealed that the average salary for men was about 8,700 higher than for women. This seemed like a considerable difference, so it was decided that a further analysis of the company salaries was warranted. 

## Loading the data


```{r load_omega_data}
omega <- read_csv(here::here("data", "omega.csv"))
glimpse(omega) # examine the data frame
```

## Relationship Salary - Gender ?

Calculation of Summary Statistics on salary by gender (mean, SD, sample size, the t-critical, the SE, the margin of error, and the low/high endpoints of a 95% condifence interval)

```{r, confint_single_valiables}
mosaic::favstats (salary ~ gender, data=omega)

#assign name to dataframe

salary_stats<- omega %>% 

#group data by gender
  
group_by(gender) %>% 

#calculate statistics 

summarise(mean_salary=mean(salary),
          
          Standard_Deviation=sd(salary),
          
          Sample_Size=n(),
          
          t_criticial=qt(0.05/2,Sample_Size-1),
          
          Standard_Error=(Standard_Deviation/sqrt(Sample_Size)),
          
          margin_of_error=Standard_Error*t_criticial,
          
          CI_high=mean_salary+margin_of_error,
          
          CI_low=mean_salary-margin_of_error)

salary_stats

```

Let's visualise these values
```{r}
ggplot(salary_stats,
       
       aes(x=reorder(gender, mean_salary), 
           y=mean_salary, 
           colour=gender)) +
  
  geom_point() +
  
  geom_errorbar(width=.2, aes(ymin=CI_low, ymax=CI_high)) + 
  
  labs(x="Gender",
       y= "Salary", 
       title= "Difference in salaries between genders: Looking at confidence intervals", 
       subtitle = "The CI's do not overlap")+ 
  
  theme_minimal()+
  
  coord_flip()+
  
  theme(legend.position = "none")+
  
  NULL
```
Conclusion

On average the salary for men and women differs by almost 10k. However, this does not mean that the difference is necessarily significant. To evaluate whether the difference is significant we can look at the confidence intervals:

The confidence Intervals do not overlap. Thus with 95% confidence we can conclude that there is a difference in population means between male and females salaries. 

Another way of determining whether the difference in salaries is significant is to test the null hypothesis= 

Mean difference in salaries is zero.

We can test this using 1) t.test() and 2) the simulation method from the infer package
 
 
 1) hypothesis testing using t.test()
```{r, hypothesis_testing}

# first we have to check whether the populations have an equal or unequal variance using the Barlett test 

salary_variance <- bartlett.test(salary ~ gender, 
        data = omega)

salary_variance

# Since the p-value is higher than 0.05, we can assume that they have equal variances.

# use t.test for equal variances
t_test <- t.test(salary~gender,
                 data = omega,
                 var.equal = TRUE)

t_test

t_test_tidy <- tidy(t_test) %>% 
  
  # Calculate difference in means, since t.test() doesn't actually do that
  
  mutate(estimate = estimate1 - estimate2) %>%
  
  # Rearrange columns
  
  select(starts_with("estimate"), everything())

t_test_tidy

```


2) Hypothesis testing using infer package
```{r}

null_distribution <- omega %>% 

# specify response variable and explanatory variable 
  specify(salary~gender) %>% 

# hypothesize H0 (use independence because we have two samples)

  hypothesize(null="independence") %>% 

# generate replicates

  generate(reps=1000, type="permute") %>% 

# calculate summary statistics
  
  calculate(stat = "diff in means", order = c("male","female"))

null_distribution

# visualize the p-value 

obs_diff_means <- omega %>%
  specify(salary~gender) %>%
  calculate(stat = "diff in means", order = c("male","female"))

actual_difference <- obs_diff_means  %>% pull()

obs_diff_means

  visualize(null_distribution, bins =10)+
  
  shade_p_value(obs_stat =obs_diff_means, direction = "both")
  
obs_diff_means %>% 
  get_p_value(obs_stat = 8696,direction = "both")

```

Conclusion: 

Both methods show us that with 95% confidence we can say that there is a statistically significant difference in salaries between men and women. 

- the t.test method shows a p value < 0.05 (the difference in means we observed is highly surprising in a world where h0 holds true) a t value whose absolute value is higher than 2, and the confidence interval does not include 0

- the infer method shows that the probability of an observed difference in means of about 8697 is very unlikely where H0=there is no difference in means between genders holds true with p value = 0  

## Relationship Experience - Gender?

At the board meeting, someone raised the issue that there was indeed a substantial difference between male and female salaries, but that this was attributable to other reasons such as differences in experience. A questionnaire send out to the 50 executives in the sample reveals that the average experience of the men is approximately 21 years, whereas the women only have about 7 years experience on average (see table below).

```{r, experience_stats}
# Summary Statistics of salary by gender
favstats (experience ~ gender, data=omega)

```
First let's have a look at the data
```{r}

# assign name to dataframe

experience_stats<- omega %>% 

# group data by gender
  
group_by(gender) %>% 

# calculate statistics 

summarise(mean_experience=mean(experience),
          
          Standard_Deviation=sd(experience),
          
          Sample_Size=n(),
          
          t_criticial=qt(0.05/2,Sample_Size-1),
          
          Standard_Error=(Standard_Deviation/sqrt(Sample_Size)),
          
          margin_of_error=Standard_Error*t_criticial,
          
          CI_high=mean_experience+margin_of_error,
          
          CI_low=mean_experience-margin_of_error)

experience_stats

```

Let's visualise this
```{r}

ggplot(experience_stats, 
       aes(x=reorder(gender, mean_experience), 
           y=mean_experience, 
           colour=gender))+
  
  geom_point() +
  
  geom_errorbar(width=.2, aes(ymin=CI_low, ymax=CI_high)) + 
  
  labs(x="Gender",
       y= "experience", 
       title="Difference in experience between genders: Looking at confidence intervals", 
       subtitle = "The CI's do not overlap") + 
  
  theme_minimal()+
  
  coord_flip()+
  
  theme(legend.position = "none")+
  
  NULL
```

Now let's do the hypothesis testing, first using t.test()
```{r}

# first we have to check whether the populations have an equal or unequal variance usng the Barlett test 

exp_variance <- bartlett.test(experience ~ gender, 
        data = omega)

exp_variance

# Since the p-value is higher than 0.05, we can assume that they have equal variances.

# use t.test for equal variances

t_test_exp <- t.test(experience~gender,
                 data = omega,
                 var.equal = TRUE)

t_test_exp

t_test_exp_tidy <- tidy(t_test_exp) %>% 
  
  # Calculate difference in means, since t.test() doesn't actually do that
  
  mutate(estimate = estimate1 - estimate2) %>%
  
  # Rearrange columns
  
  select(starts_with("estimate"), everything())

t_test_exp_tidy
```
Now with the infer package
```{r}

null_distribution_exp <- omega %>% 

# specify response variable and explanatory variable 
  specify(experience~gender) %>% 

# hypothesize H0 (use independence because we have only one dataset)

  hypothesize(null="independence") %>% 

# generate replicates

  generate(reps=1000, type="permute") %>% 

# calculate summary statistics
  
  calculate(stat = "diff in means", order = c("male","female"))

null_distribution

# Visualize p value 

obs_diff_means_exp <- omega %>%
  
  specify(experience~gender) %>%
  
  calculate(stat = "diff in means", order = c("male","female"))

obs_diff_means_exp
  
  visualize(null_distribution_exp, bins =10)+
  
  shade_p_value(obs_stat =obs_diff_means_exp, direction = "both")
  
obs_diff_means %>% 
  get_p_value(obs_stat = obs_diff_means_exp, direction = "both")
  
```
Conclusion: 

Both methods show us that there is a statistically significant difference in experience between men and women

- summary statistics show that the confidence intervals do not overlap 

- the t.test method shows a p value < 0.05, a t value whose absolute value is higher than 2, and the confidence interval does not include 0 

- the infer method shows that the probability of an observed difference in means of about 14 years is very unlikely (p value = 0) when H0=there is no difference in experience between genders holds true 

There is still a significant difference in female and male salaries, however, this additional analysis shows that this difference is partly caused by the higher years of experience of the male executives. Overall, it weakens the argument that the difference in pay is solely due to discrimination. 


## Relationship Salary - Experience ?

Let's have a closer look at the relationship between experience in salary using a scatter plot

```{r, salary_exp_scatter}
# create plot from data omega where x axis shows experience, and y axis shows salary

ggplot(omega, aes(x = experience, y=salary,color=gender))+
  
# use geom_point to create a scatterplot
  
  geom_point()+
  
# add a theme and labels
  
  theme_economist_white()+
  
  labs(
    title = "Strong positive relationship between experience and salary",
    subtitle = "Relationship between salary and years of experience",
    x ="Years of Experience",
    y ="Salary")+
  
  # remove legend 
  
  theme(legend.position = "none")+
  
  NULL
```


## Check correlations between the data
You can use `GGally:ggpairs()` to create a scatterplot and correlation matrix. Essentially, we change the order our variables will appear in and have the dependent variable (Y), salary, as last in our list. We then pipe the dataframe to `ggpairs()` with `aes` arguments to colour by `gender` and make ths plots somewhat transparent (`alpha  = 0.3`).

```{r, ggpairs}
omega %>% 
  select(gender, experience, salary) %>% #order variables they will appear in ggpairs()
  ggpairs(aes(colour=gender, alpha = 0.3))+
  theme_bw()
```

Conclusions from scatterplot

Overall, there is a positive relationship between experience in salary. However, this experience appears to not be completely linear. Between 0 and 10 years salary increases strongly as experience increases, while this relationship flattens off beyond 10 years.  
