---
title: "Covid 19 Analysis"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure sizes
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(maps)
library(viridis)
```

# CDC COVID-19 Public Use Data

```{r, cache=TRUE}
# file contains 11 variables and 3.66m rows and is well over 380Mb. 
# It will take time to download

# URL link to CDC to download data
url <- "https://data.cdc.gov/api/views/vbim-akqf/rows.csv?accessType=DOWNLOAD"

covid_data <- vroom::vroom(url)%>% # If vroom::vroom(url) doesn't work, use read_csv(url)
  clean_names()


```

Given the data we have, I will produce two graphs that show death % rate:

1. by age group, sex, and whether the patient had co-morbidities or not
1. by age group, sex, and whether the patient was admited to Intensive Care Unit (ICU) or not.


```{r}
skim (covid_data)
```
By Age group, sex, and morbidity
``` {r}
covid_data_clean_1 <- covid_data %>% 
filter(death_yn!="Missing") %>% 
  filter(death_yn!="Unknown") %>% 
  filter(age_group!="Unknown") %>% 
  filter(sex!="Unknown") %>%
  filter(sex!="Missing") %>% 
  filter(medcond_yn!="Missing") %>%
  filter(medcond_yn!="Unknown") %>%
  filter(sex!="Other")
```

```{r}
         
covid_morb<- covid_data_clean_1 %>% 
  group_by(age_group,death_yn,sex,medcond_yn) %>%
  summarise(count=n()) %>% 
  pivot_wider(names_from = death_yn, values_from = count) %>% 
  mutate(death_rate=(Yes/(Yes+No))) %>% 
  mutate(death_rate_perc=paste(round(100*death_rate, 1), "%", sep=""))

covid_morb

covid_morb$medcond_yn_f = factor(covid_morb$medcond_yn, levels=c('Yes','No'))

morb.labs <- c("Without comorbidities", "With comorbidities")
names(morb.labs) <- c(medcond_yn_f="No", medcond_yn_f="Yes")

ggplot(covid_morb, aes(x=death_rate,y=age_group))+
  geom_col(alpha=0.8,fill="#6C7EA2")+
  facet_grid(medcond_yn_f~sex, 
               ,labeller=labeller(medcond_yn_f=morb.labs))+
  labs(title="Covid death rate by age group, sex and presence of co-morbidities",x="",y="")+
  scale_x_continuous(labels=scales::percent)+
  geom_text(aes(label=death_rate_perc),nudge_x = 0.07, size=3)+
  coord_fixed(ratio = 0.04)+
   theme(strip.background = element_rect(
     color="black", fill="grey", linetype="solid"), panel.background = element_rect(fill = "white"),panel.grid=element_line(size=0.5, colour="#F1F1EB"), panel.border = element_rect(fill=NA))
```
By age group, sex and admissions to ICU 

``` {r}
covid_data_clean_2 <- covid_data %>% 
  filter(death_yn!="Missing") %>% 
  filter(death_yn!="Unknown") %>% 
  filter(age_group!="Unknown") %>% 
  filter(sex!="Unknown") %>%
  filter(sex!="Missing") %>% 
  filter(icu_yn!="Missing") %>%
  filter(icu_yn!="Unknown") %>%
  filter(medcond_yn!="Missing") %>%
  filter(medcond_yn!="Unknown") %>%
  filter(sex!="Other")
         
covid_icu<- covid_data_clean_2 %>% 
  group_by(age_group,death_yn,sex,icu_yn) %>%
  summarise(count=n()) %>% 
  pivot_wider(names_from = death_yn, values_from = count) %>% 
  mutate(death_rate=(Yes/(Yes+No))) %>% 
  mutate(death_rate_perc=paste(round(100*death_rate, 1), "%", sep=""))

covid_icu

covid_icu$icu_yn_f = factor(covid_icu$icu_yn, levels=c('Yes','No'))

icu.labs <- c("Not Admitted to ICU", "Admitted to ICU")
names(icu.labs) <- c(icu_yn_f="No", icu_yn_f="Yes")

ggplot(covid_icu, aes(x=death_rate,y=age_group))+
  geom_col(alpha=0.8,fill="#FF5B33")+
  facet_grid(icu_yn_f~sex, 
               , labeller=labeller(icu_yn_f=icu.labs))+
  labs(title="Covid death rate by gender, age and ICU admittance",x="",y="")+
  scale_x_continuous(labels=scales::percent)+
  geom_text(aes(label=death_rate_perc),nudge_x = 0.07, size=3)+
  coord_fixed(ratio = 0.04)+
  theme(strip.background = element_rect(
     color="black", fill="grey", linetype="solid"), panel.background = element_rect(fill = "white"),panel.grid=element_line(size=0.5, colour="#F1F1EB"), panel.border = element_rect(fill=NA))
```


# Covid cases in other countries 

Given that this data was collected by the CDC it only provides information on the covid epidemic in the US. As an european I would like to have a closer look at Covid in Europe.

```{r}
#download the data
covid_more<- read_csv("~/desktop/Hanna-website/data/coviddata.csv")

#have a look at the data
skim(covid_more)
```

First let's compare the number of total cases in a few European countries
```{r}

covid_europe <- covid_more %>%
  
#filter data for countries
  filter(location %in% c("United Kingdom", "Germany","France","Sweden", "Switzerland"))

covid_europe
```
```{r}
ggplot(covid_europe, aes(x=date, y=total_cases, fill=location))+
  
#choose graph type
  geom_col()+
  
#facet by country
  facet_wrap(~location)+
  
#add labels
  labs(title = "France and the United Kingdom have most covid cases
       ", 
        x="
       Time", 
       y="Total Cases
       ")+
  
#choose theme and remove legend
  theme_economist()+
  theme(legend.position = "none")+
  NULL
```
Now that we know that the UK and France have the highest total number of covid cases. Let's have a look at the measures these countries have put in place in response. 
```{r}
ggplot(covid_europe, aes(x=date, y=stringency_index, color=location))+
  
#choose graph type
   geom_line()+

# add labels
  labs(title = "France had the strictest covid measures
       ", x="
       Time", 
       y="Stringency Index
       ")+

# choose theme & legend position & title
  theme_economist()+
  theme(legend.position = "right")+
  theme(legend.title = element_blank())+
  NULL
```
As you can see from the Graph the UK and France also have the strictest covid measures as measured by the stringency index. 

Now let's have a look at whether strict measures have an effect on daily new cases.
```{r}
ggplot(covid_europe, aes(x=date, y=new_cases, color=stringency_index))+
  
#choose graph type
  geom_point()+

# choose color scale
  scale_color_gradient(low="blue", high="red")+
  
# add labels
  labs(title = "Strict measures lead to lower case numbers
       ", x="
       Time", 
       y="New Cases
       ")+
  
#change legend title
  guides(color=guide_legend("Stringency Index"))+
  
#facet wrap by country
  facet_wrap(~location)+
  
#choose theme
  theme_economist()+
  
  NULL
```
Now onto what we really care about: is there any country in this world where you do not have to adhere to strict covid measures? Well I have created a map. 

```{r}
#Transform data
covid_measures <- covid_more %>%

#select only relevant variables
  select(iso_code,location,date,stringency_index) %>%

#rename some columns for join with 
  rename(
    region = location,
   Stringency = stringency_index
    ) %>% 
#rename United States to USA for left join
  mutate(
    region = ifelse(region == "United States", "USA", region)
    ) %>% 

#filter for a recent date (i choose this because there were only few stringency index values available for more recent data)
  filter(date=="2020-10-01")
```

```{r}
#create world map data frame
world_map <- map_data("world")

#join data frames
covid_measures_map <- left_join(covid_measures, world_map, by = "region")

```

```{r}
#plot 
ggplot(covid_measures_map, aes(map_id = region, fill = Stringency))+

#choose geom_map
  geom_map(map = covid_measures_map,  color = "white")+

#adapt limits of x and y scale
  expand_limits(x = life.exp.map$long, y = life.exp.map$lat)+

#select scale fill
  scale_fill_viridis_c(option = "C")+
  
#add labels
  labs(title= "Most countries currently have intermediate measures in place",subtitle = "Current Stringency level of covid measures in the world", x="",y="")+
  NULL
```


