---
title: "About me"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

## Short Summary

 * Currently reading for the Masters in Financial Analysis at London Business School 
 * Previously worked in M&A and Private Equity 
 * Recently started working with R

## Education

```{r}
library(tibble)
library(kableExtra)

degree <-c("MSc in Financial Analysis","BSc in International Business Administration","Global BBA (Semester Abroad)", "International Baccalaureate")

years <- c("2020-2022","2016-2019","2018", "2013-2016")

institution <- c("London Business School", "WHU - Otto Beisheim School of Management", "Essec Business School", "King Edward's School")

location <- c("London, UK", "Vallendar, GER", "Paris, FR", "Witley, UK")

education_frame <- data.frame(Date=years, Location=location, Institution=institution, Degree=degree)

kable(education_frame)

```


## Work Experience
```{r}
Dates <-c("2020","2019/2020", "2019","2018","2016")

Company <- c("Rothschild & Co","Silver Investment Partners","Enpal", "Evonik Nippon Aerosil", "Pharmalex")

Position <- c("Intern, M&A", "Intern, Investment Team","Intern","Intern, Marketing Team", "Intern, M&A Department")

Place <- c("Frankfurt, GER", "Königstein, GER", "Berlin, GER", "Tokyo, JP", "Friedrichsdorf, GER")
 
jobs_frame <- data.frame(Date=Dates, Location=Place, Company=Company, Position=Position)

kable(jobs_frame)

```

## Extra-curricular activities 

* LBS Women in Business Club IB/PE/AM Lead: Empowering women in Finance through building networks and inspiring the next generation of leaders
* Member of WHU's National Model United Nations Club: Participated in conferences in Hamburg and New York with aim of defending a country's position in a plenum and collaborating with other countries in developing resolutions on aftermath of Syrian civil war and counter-terrorism efforts
* Workshops & Competitions: Morgan Stanley Future Female Traders Workshop (2020), BCG Case Workshop on Digitalisation and Social Impact (2019)
* Hobbies: Tennis (14 years), skiing (18 years), reading (recent book: Prisoners of Geography)


